### 2020-07-09 abstract.tex
Katy's second review edits

Molten salt reactors are a class of advanced nuclear reactors that promise
numerous improvements over the current fleet of largely light-water reactors.
As the world continues its transition towards low-carbon electricity
generation to combat climate change, molten salt reactors are potential
options in the near-term future for replacing fossil fuel and ageing nuclear
power plants. At the current state of development, molten salt reactors still
require extensive research to become viable. This thesis presents latest
developments in Moltres, a simulation tool for molten salt reactors. These new
developments are: support for coupling the incompressible Navier-Stokes and
the delayed neutron precursor looping systems, and a new decay heat model for
simulating decay heat from fission products at steady state and during
transients. This work demonstrates these capabilities in conjunction with
existing capabilities through multiphysics simulations of the Molten Salt Fast
Reactor concept. This work first verifies the six-group neutron diffusion
results from Moltres against continuous-energy Monte Carlo neutron transport
results from Serpent 2. The multiplication factors $k_{\text{eff}}$, delayed
neutron fractions $\beta$, temperature reactivity coefficient $\alpha_T$, and
the six-group neutron energy spectra from Moltres agreed with the high
fidelity simulation results from Serpent 2. The $k_{\text{eff}}$ values have
small discrepancies on the order of 100 pcm, which is smaller than some
reported literature values. The decay heat model shows an expected flattening
of the temperature distribution due to the movement and dispersion of the
decay heat precursors throughout the primary coolant loop. This work also
demonstrates and verifies steady state and
transient multiphysics simulations of the Molten Salt Fast Reactor. The
transient scenarios under study are unprotected instances of reactivity
insertion, loss of heat sink, loss of flow, and pump overspeed. The steady
state and transient results are verified against results from another paper
that presented results for the same cases. The steady-state temperature and
velocity distributions, and the peak neutron flux showed good agreement with
the literature results. Minor differences in the delayed neutron precursor
distribution and the in-core delayed neutron fraction were explainable with
the differences in the handling of turbulence in the models. In three of the
transient results (reactivity insertion, loss of heat sink, and pump
overspeed), Moltres reproduced the expected magnitude and pattern of the
reactor response to these transient initiators.
The loss of flow results showed greater discrepancies that are attributed
to differences in the fluid dynamics modeling in Moltres and the other models.
Through the verification studies, this work has also identified avenues
for further Moltres software development.

>>>>

Molten salt reactors is a class of advanced nuclear reactors that promise
numerous improvements over the current fleet of largely light-water reactors.
As the world continues its transition towards low-carbon electricity
generation to combat climate change, molten salt reactors is potential
options in the near-term future for replacing fossil fuel and ageing nuclear
power plants. At the current state of development, molten salt reactors still
require extensive research to become viable. This thesis presents the latest
developments in Moltres, a simulation tool for molten salt reactors. These new
developments are: support for coupling the incompressible Navier-Stokes and
the delayed neutron precursor looping systems, and a model for
simulating decay heat from fission products at steady state and during
transients. This work demonstrates these capabilities through multiphysics
simulations of the Molten Salt Fast
Reactor concept. This work first verifies the six-group neutron diffusion
results from Moltres against continuous-energy Monte Carlo neutron transport
results from Serpent 2. The multiplication factors $k_{\text{eff}}$, delayed
neutron fractions $\beta$, temperature reactivity coefficient $\alpha_T$, and
the six-group neutron energy spectra from Moltres agreed with the high
fidelity simulation results from Serpent 2. The $k_{\text{eff}}$ values have
small discrepancies on the order of 100 pcm, which is smaller than the
$-256.7$ pcm discrepancy reported in literature with the same six-group
neutron diffusion approach. The decay heat model shows an expected flattening
of the temperature distribution due to the movement and dispersion of the
decay heat precursors throughout the primary coolant loop. This work also
demonstrates and verifies steady state and
transient multiphysics simulations of the Molten Salt Fast Reactor. The
transient scenarios under study were unprotected instances of reactivity
insertion, loss of heat sink, loss of flow, and pump overspeed. The steady
state and transient results were verified against results from another paper
that presented results for the same cases. The steady-state temperature and
velocity distributions, and the peak neutron flux showed good agreement with
the literature results. Minor differences in the delayed neutron precursor
distribution and the in-core delayed neutron fraction were explainable with
the differences in the handling of turbulence in the models. In three of the
transient results (reactivity insertion, loss of heat sink, and pump
overspeed), Moltres reproduced the expected magnitude and pattern of the
reactor response to these transient initiators.
The loss of flow results showed greater discrepancies that are attributed
to differences in the fluid dynamics modeling in Moltres and the other models.
Through the verification studies, this work has also identified avenues
for further Moltres software development.

### 2020-07-09 introduction.tex
Katy's second review edits

Greenhouse gas emission from human activities is the main cause of climate
change, a global phenomenon characterized by shifting weather patterns and
rising temperatures. The concentration of CO$_2$, a greenhouse gas, in the
atmosphere has been rising steadily throughout modern human history
\cite{us_department_of_commerce_global_2020}. Climate change has dire
consequences on human health and safety due to extreme weather events and the
overall impact on food production \cite{mcmichael_global_2004}. As a result,
urgent measures are necessary to limit greenhouse gas emissions, namely CO$_2$
emissions from electricity generation.

Electricity generation from burning fossil fuels represents the
greatest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019});
replacing it with low-carbon
alternatives would curb a significant fraction of emissions. Nuclear power is
a viable low-carbon replacement for burning fossil fuels.
Other low-carbon energy sources include solar-, wind-, and hydro-power. Each
energy source has its own set of advantages and disadvantages; nuclear power's
advantage over renewable energy sources is that it provides consistent
base-load power independent of weather and geographical location
\cite{petti_future_2018}. Employing a diverse mix of these energy sources
ensures energy security and reliability in our transition
towards a low-carbon future \cite{petti_future_2018}.

@@

However, the liquid fuel form also brings about novel computational
challenges in simulating the transient behavior of \glspl{MSR}; the
interactions between neutronics and thermal-hydraulics are stronger due
to greater fuel material expansion. Furthermore, \glspl{DNP} in \glspl{MSR}
can flow freely within the primary coolant
loop as opposed to being held in place in a solid fuel matrix. Therefore,
the choice of coupling methods for each set of physics requires careful
consideration.

Most reactor analysis applications are usually reactor-specific by
design such as PARCS and TRACE for \glspl{LWR}, and SAS4A/SASSYS-1 for
liquid metal cooled reactors. Thus, these applications would disregard
\gls{MSR}-specific phenomena and are inappropriate for \gls{MSR}
analysis without any modifications to the source code. Some research efforts
do focus on adapting these applications for \gls{MSR} analysis. Examples
include the coupling of modified versions of TRACE and PARCS
\cite{pettersen_coupled_2016}, and the development of VERA-MSR from the
integrated \gls{LWR} simulation tool VERA \cite{graham_development_2019}.
Others developed their \gls{MSR} simulation tools from general
multiphysics or \gls{CFD} applications such as COMSOL
\cite{fiorina_modelling_2014} and OpenFOAM \cite{aufiero_development_2014}.

Moltres is an open source MSR simulation tool that falls under the latter
category \cite{lindsay_introduction_2018}. It is an application built in the
\gls{MOOSE} \cite{gaston_physics-based_2015} parallel finite element
framework. Lindsay et al. first presented the tool in 2017 and demonstrated
its capabilities by simulating 2-D and 3-D models of the \gls{MSRE}
\cite{lindsay_introduction_2018}. The results showed good qualitative
agreement with the original design calculations by \gls{MSRE} researchers at
\gls{ORNL}. There has since been new developments in Moltres allowing for
more accurate \gls{MSR} simulations.

@@

This thesis demonstrates latest capabilities of Moltres
\cite{lindsay_introduction_2018}, an open source MSR simulation tool.
In particular, this thesis presents two more recent
developments in Moltres, namely fully integrating \gls{MOOSE}'s incompressible
Navier-Stokes module into Moltres, and introducing a
decay heat model.
The main objective of this thesis is to verify Moltres'
latest capabilities in modeling multiphysics, steady-state, and transient
behavior of fast-spectrum \glspl{MSR} through the study of the \gls{MSFR}
concept. Code-to-code verification is an important exercise in software
development for ensuring that the application produces accurate and reliable
results. This thesis covers the \gls{MSFR} concept mainly because it has been
studied extensively with readily available literature data to verify our
results against. The
steady-state and transient results also feature a significant degree of
physics interaction between the neutronics and thermal-hydraulics. This
present work will first present a verification of Moltres' \gls{MSFR}
neutronics results against Serpent 2, followed by a verification of
the coupled neutronics/thermal-hydraulics steady-state and transient accident
results against results from the Polimi and TUDelft models published by
Fiorina et al. \cite{fiorina_modelling_2014}. The
secondary objective is to identify areas of improvement in Moltres for future
research.

@@

The outline of this thesis is as follows. Chapter 2 discusses the history and
features of \glspl{MSR}, and a literature review of existing \gls{MSR}
simulation tools. The chapter also covers the \gls{MSFR} concept in greater
detail. Chapter 3 details the software and the general modeling
approach for generating the results in this thesis. Chapter 4 provides a
neutronics assessment by comparing key neutronics parameters from Moltres'
eigenvalue calculations to Serpent's Monte Carlo calculations. Chapter 5
presents steady-state results of coupled neutronics/thermal-hydraulics
\gls{MSFR} simulations in Moltres. Chapter 6 presents transient accident
simulation results for unprotected reactivity insertions, unprotected loss of
heat sink, unprotected loss of flow, and unprotected pump overspeed,
respectively. Lastly, Chapter 7 summarizes the key findings in this thesis
and posits some potential avenues for future work.

>>>>

Greenhouse gas emission from human activities is the main cause of climate
change which has dire
consequences on human health and safety due to extreme weather events and the
overall impact on food production \cite{mcmichael_global_2004}.
Electricity generation from burning fossil fuels represents the
greatest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019});
replacing it with low-carbon
alternatives would curb a significant fraction of emissions. Nuclear power is
a viable low-carbon replacement for burning fossil fuels and it provides
consistent base-load power independent of weather and geographical location
\cite{petti_future_2018}. Furthermore, employing a diverse mix of nuclear
power and renewable sources ensures energy security and reliability in our
transition towards a low-carbon future \cite{petti_future_2018}.

@@

However, the liquid fuel form also brings about novel computational
challenges in simulating the transient behavior of \glspl{MSR}; the
interactions between neutronics and thermal-hydraulics are stronger due
to greater fuel material expansion. Furthermore, fissile material and
\glspl{DNP} in \glspl{MSR} can flow freely within the primary coolant
loop as opposed to being held in place in a solid fuel matrix. Therefore,
the choice of coupling methods for each set of physics requires careful
consideration.

Most reactor analysis applications are usually reactor-specific by
design such as TRACE for \glspl{LWR}, and SAS4A/SASSYS-1 for
liquid metal cooled reactors. Thus, these applications would disregard
\gls{MSR}-specific phenomena and are inappropriate for \gls{MSR}
analysis without any modifications to the source code. Some research efforts
do focus on adapting these applications for \gls{MSR} analysis. Examples
include the coupling of modified versions of TRACE and PARCS
\cite{pettersen_coupled_2016}, and the development of VERA-MSR from the
integrated \gls{LWR} simulation tool VERA \cite{graham_development_2019}.
Others developed their \gls{MSR} simulation tools from general
multiphysics or \gls{CFD} applications such as COMSOL
\cite{fiorina_modelling_2014} and OpenFOAM \cite{aufiero_development_2014}.

Moltres is an open source MSR simulation tool that falls under the latter
category \cite{lindsay_introduction_2018}. It is an application built in the
\gls{MOOSE} \cite{gaston_physics-based_2015} parallel finite element
framework. Lindsay et al. first presented the tool in 2017 and demonstrated
its capabilities by simulating 2-D and 3-D models of the \gls{MSRE}
\cite{lindsay_introduction_2018}. The results showed good qualitative
agreement with the original design calculations by \gls{MSRE} researchers at
\gls{ORNL}. This thesis presents some of the new developments in Moltres
allowing for more complex and accurate \gls{MSR} simulations.

@@

This thesis demonstrates latest capabilities of Moltres
\cite{lindsay_introduction_2018}.
In particular, this thesis presents two more recent
developments in Moltres, namely fully integrating \gls{MOOSE}'s incompressible
Navier-Stokes module into Moltres, and introducing a
decay heat model.
The main objective of this thesis is to verify Moltres'
latest capabilities in modeling multiphysics, steady-state, and transient
behavior of fast-spectrum \glspl{MSR} through the study of the \gls{MSFR}
concept. Code-to-code verification is an important exercise in software
development for ensuring that the application produces accurate and reliable
results. This thesis covers the \gls{MSFR} concept mainly because it has been
studied extensively with readily available data in the literature to verify
our results against. The \gls{MSFR} design also features interesting flow
patterns that greatly affect the steady-state and transient behavior. This
present work will first present a verification of Moltres' \gls{MSFR}
diffusion neutronics against the Monte Carlo neutron tranport software
Serpent 2, followed by a verification of
the coupled neutronics/thermal-hydraulics steady-state and accident transient
results against results from the Polimi and TUDelft models published by
Fiorina et al. \cite{fiorina_modelling_2014}. The
secondary objective is to identify areas of improvement in Moltres for future
research.

@@

The outline of this thesis is as follows. Chapter 2 discusses the history and
features of \glspl{MSR}, and a literature review of existing \gls{MSR}
simulation tools. The chapter also covers the \gls{MSFR} concept in greater
detail. Chapter 3 details the software and the general modeling
approach for generating the results in this thesis. Chapter 4 provides a
neutronics assessment by comparing key neutronics parameters from Moltres'
eigenvalue calculations to Serpent's Monte Carlo calculations. Chapter 5
presents steady-state results of coupled neutronics/thermal-hydraulics
\gls{MSFR} simulations in Moltres. Chapter 6 presents accident transient
simulation results for unprotected reactivity insertions, unprotected loss of
heat sink, unprotected loss of flow, and unprotected pump overspeed. Lastly,
Chapter 7 summarizes the key findings in this thesis
and posits some potential avenues for future work.
